\documentclass[12pt,a4paper]{book}\usepackage{array}\usepackage{tabu}\usepackage{multirow}\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}\usepackage{multirow}\usepackage[utf8]{inputenc}\usepackage[T1]{fontenc}\usepackage{lmodern}\usepackage{hyperref}\usepackage{graphicx}\usepackage[english, frenchb]{babel}\usepackage{listings}	\lstdefinestyle{customstyle}{basicstyle=\footnotesize,breakatwhitespace=false,breaklines=true,captionpos=b,keepspaces=true,tabsize=4,frame=single,moredelim=[is][\underbar]{_}{_}}\lstset{style=customstyle}\usepackage{menukeys}\usepackage{comment}\begin{document}
\chapter{Git server}
\section{Why you would use Git}
For some reasons, you need to keep several versions of document so you make 
copies -- can be long and space consumer -- but it still difficult to get the 
last version when more than one person works on one document.
\\\\
Git is a versionning tool which can do this work for you and in addition keep 
traces of who modified what on each document on a project. This tool will be not  
intrusive and can be used at every moments of a projet -- even if it is allready 
started -- it ensure that you have a project which is \og{}good\fg{} and 
up-to-date version.
\\\\
Largely used in programming projects it can be usefull for any types of projects 
which contains files readable by a simple text editor. You can get an idea of what 
is git by browsing projects on \href{http://github.com}{github.com} which is the 
most famous git projects host because it is free for public (open source) projets.

\section{Install Git and add projects}
To visualize git projects on web interface the first thing is need to have git 
project. In order to create our own projet we will download an existing one 
to have a real history and some things to visualize. 
\\\\
I choose to use an open source project -- Firefox OS\footnote{Firefox OS is the 
operating system developed by the Mozilla foundation for smartphones. More 
informations on \href{https://www.mozilla.org/en-US/firefox/os/}{www.mozilla.org/
en-US/firefox/os/}, sources availables on \href{https://github.com/mozilla/
firefox-ios}{github.com/mozilla/firefox-ios}} -- because it is quite fast to 
download and have some folders with enough commits to be interesting.

\begin{lstlisting}[language=bash,caption=Install git and add projects]
# Install git package
$ pacman -S git

# Create repositories(projects) directory
$ mkdir /srv/git

# Get an existing project on github to test our web 
# interface, Firefox OS in this example.
$ cd /srv/git
$ git clone https://github.com/mozilla/firefox-ios.git
\end{lstlisting}

There are a few directories that are most used for repositories but I retained two 
of them: \texttt{/home/git/} and \texttt{/srv/git/}. 
\\\\
In ArchLinux, websites and ftp are by default in \emph{/srv} and \emph{/home/git} 
is the home directory for \emph{git} user -- which is not a physic user -- so to 
keep a kind of consistency -- in my opinion -- I prefer the second alternative. 

\section{Host repositories}
\subsection{Create a new project}
Host a git project is quite easy because it is really close from creating a local 
git repository. You just must think about adding \emph{--bare} option to specify 
that your project is not a working copy. 
\begin{lstlisting}[language=bash,caption=Create a repository on your server]
# Let's create a new git project called "project"
# Naming convention: add .git at the end of your project name
# to specify it is not a working copy
#
$ mkdir /srv/git/project.git
$ cd /srv/git/project.git
$ git init --bare # --bare: not a working copy
\end{lstlisting}

On a remote machine on which you want to work, clone the project and do a first 
commit on your project.
\begin{lstlisting}[language=bash,caption=Clone repository to work]
# Clone project
# username is your username for local account on the server 
$ git clone username@RPiIP:/srv/git/project.git

# Create a new file and commit
$ cd project
# New file called README.md with "New project" text inside
$ echo "New project" > README.md 
$ git commit -am"First commit"
$ git push origin master # First time, after just "git push"
\end{lstlisting}

\subsection{Use an existing project}
If you want to host a git project which allready it is possible to do. There two 
possible situations:
\begin{itemize}
	\item The project is allready hosted by someone else or websites like GitHub 
		  or BitBucket.
	\item It is a local project so hosted locally
\end{itemize}

\subsubsection{Allready hosted}
As you can imagine, the first situation is the easier to solve, you just have to 
clone the repository from the server with a specific option.

\begin{lstlisting}[language=bash]
$ git clone username@RPiIP:/srv/git/project.git --bare
\end{lstlisting}

Note that the folder you will get will be named \emph{project.git} so it is not a 
working copy. In addition, notice that is the same \emph{--bare} option as when 
you create a new repository on the server.

\subsubsection{Local project}
The first thing to do is retrieve the local project on the server. There are many 
ways to it, I advice to use FTP -- if you setup an FTP -- but for others here is 
a simple way if you have a linux system.

\begin{lstlisting}[language=bash]
# Your project is named project
$ scp username@RPiIP:/srv/git /local/path/to/project
\end{lstlisting}

With the working copy on your server, the last thing to do is clone the repository 
as \emph{bare} locally and remove the working copy.

\begin{lstlisting}[language=bash][Host repository from local copy]
# Extract project informations
$ git clone --bare /srv/git/projet /srv/git/project.git
$ rm -r /srv/git/project
\end{lstlisting}
% Change remote ?
	
% Different from clone apparently
\begin{comment}
	\begin{lstlisting}[language=bash][Host repository from local copy]
# Extract project informations
$ mv /srv/git/project/.git /srv/git/project.git
$ rm -r /srv/git/project

# Declare repository as bare
$ cd /sr	v/git/project.git
$ git config --bool core.bare true
	\end{lstlisting}
\end{comment}
\section{Web application for viewing}
\subsection{GitWeb: Build-in solution}
Gitweb is the web interface offered by git when you install it. Based on cgi 
technology, it is a bit slow but very refined so easy to use and understand.
\\\\
You can see a live usage of git web at \href{http://repo.or.cz/w?a=project\_list}
{repo.or.cz/w?a=project\_list}, browse projets, commit and all things you want and 
before abandoning GitWeb keep in mind that we will improve a little bit the visual.

\subsubsection{Installation and configuration}
\begin{lstlisting}[language=bash,caption=Gitweb setup]
# Gitweb is a cgi script, you need to install 
# some packages
$ pacman -S fcgiwrap spawn-fcgi

# Update php configuration to accept .cgi files
$ nano /etc/php/php-fpm.conf
# Uncomment and add .cgi
security.limit_extensions = .php .php3 .php4 .php5 .cgi

# Set your repositories folder in gitweb.conf
# Look at the next example
$ nano /etc/gitweb.conf  # Is created of not exist
\end{lstlisting}

\lstset{language=bash,caption=gitweb.conf file}
\lstinputlisting{code/gitweb.conf}

\subsubsection{Configure nginx and install cgi tools}
According to the ArchLinux wiki, it is necessary to change the \emph{fcgiwrap.service} 
file because it uses spawn-fcgi so trust the wiki. Note you have to restart 
fcgiwrap service after your updates.

\lstset{language=bash,caption=fcgiwrap.service from wiki.archlinux.org/index.php/gitweb}
\lstinputlisting{code/fcgiwrap.service}

The standard subdomain file presented before will not work in this case because 
gitweb does not contains php scripts but only one cgi script.

\lstset{language=bash,caption=Gitweb configuration for nginx}
\lstinputlisting{code/gitweb}

\subsubsection{Finishing touch}
After that, you will be able to access to your gitweb interface with the domain 
you defined -- \texttt{gitweb.local.pi} in the example -- but if do not see any 
projects although you set some of them in the directory you specified in 
\emph{gitweb.conf}, try to launch git daemon.

\begin{lstlisting}[language=bash,caption=]
$ /usr/lib/git-core/git-daemon --inetd --export-all --base-path=/srv/git
\end{lstlisting}

\mbox{}\\\\
The last thing you can do is replace the default theme the one developed by 
kogakure\footnote{More informations on his web site: \href{http://kogakure.github
.io/gitweb-theme/}{kogakure.github.io/gitweb-theme/}} which is a little bit more 
appealing.

\begin{lstlisting}[language=bash,caption=Install kogakure gitweb theme]
# Save default theme
$ cp -r /usr/share/gitweb/static/ /usr/share/gitweb/static.old

# Retrieve kogakure theme, copy it in right folder
# and delete unused files
$ git clone https://github.com/kogakure/gitweb-theme.git
$ cp gitweb-theme/git* /usr/share/gitweb/static
$ rm -r gitweb-theme
\end{lstlisting}

\subsection{GitList: Simple GitHub clone}

If you are looking for a git web interface which look like GitHub -- with less 
side features -- GitList is probably a good alternative for you. 
\\\\
This project is hosted by GitHub at \href{https://github.com/klaussilveira/gitlist}
{github.com/klaussilveira/gitlist} but as said in the README file, do not clone 
the repository and prefer download the packaged version on the GitList website at 
\href{http://gitlist.org}{gitlist.org}.

\begin{lstlisting}[language=bash,caption=Download GitList]
# Download
$ cd /usr/share/webapps
$ wget https://s3.amazonaws.com/gitlist/gitlist-0.5.0.tar.gz

# Extract and remove unused files
$ tar -xvf gitlist-0.5.0.tar.gz
$ rm gitlist-0.5.0.tar.gz
\end{lstlisting}

Let's declare a new subdomain for GitList, a sample is available in it's directory 
-- \texttt{/usr/share/webapps/gitlist/INSTALL.md} -- but here I will propose you 
a shorter solution. Note you have to use at least one of these two sample because 
GitList uses url rewriting.

\lstset{language=bash,caption=GitList configuration for nginx}
\lstinputlisting{code/gitlist}

Now you can access to GitList interface but you may have some errors messages and 
your are forced to fix these troubles.

\begin{lstlisting}[language=bash,caption=Fix GitList troubles]
# Message 1
# The "/usr/share/webapps/gitlist/cache" folder must be writable 
# for GitList to run.
#
$ mkdir /usr/share/webapps/gitlist/cache # If not exist
$ chown http:http /usr/share/webapps/gitlist/cache

# Message 2
# Please, create the config.ini file.
$ cd /usr/share/webapps/gitlist/
$ mv config.ini-example config.ini

# Message 3
# Please, edit the config file and provide your repositories directory
# Solution: set repositories[] = '/srv/git/'
#
$ nano /usr/share/webapps/gitlist/config.ini
\end{lstlisting}

\section{Gitolite: manage projects}
Soon in libraries

\end{document}
